---
title: "Writing Special Cell Types with writexl"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing Special Cell Types with writexl}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(writexl)
```

The `writexl` package exports plain data frame values (numbers, strings,
booleans, dates) automatically. For anything more specialised — formulas,
hyperlinks, or mixed-type columns — the package provides the `xl_cell_general`
family of helpers.

## 1. Formulas

Use `xl_formula()` to mark character strings as Excel formulas. Each formula
**must** start with `"="`. A length-1 formula is automatically recycled to fill
the whole column.

```{r formulas}
df <- data.frame(
  name    = c("UCLA", "Berkeley", "Caltech"),
  founded = c(1919, 1868, 1891)
)
df$age <- xl_formula('=(YEAR(TODAY()) - INDIRECT("B" & ROW()))')
```

You can also supply a **pre-calculated result** so that xlsx readers that do
not recalculate on open still show a sensible value:

```{r formula-precalc}
cells <- c(
  xl_cell_general(value = 55.0, formula = "=SUM(B2:B4)"),
  xl_cell_general(value = 10.5, formula = "=AVERAGE(B2:B4)")
)
df2 <- data.frame(label = c("Sum", "Mean"))
df2$result <- cells
```

When `value` is numeric the package uses `worksheet_write_formula_num()` so
the stored value is visible even without recalculation. When `value` is a
character string, `worksheet_write_formula_str()` is used instead.

## 2. Hyperlinks

Use `xl_hyperlink()` for clickable URLs. Optionally supply a `name` vector
for the display text shown in the cell:

```{r hyperlinks}
df3 <- data.frame(
  name = c("UCLA", "Berkeley"),
  url  = xl_hyperlink(
    c("http://www.ucla.edu", "http://www.berkeley.edu"),
    name = "homepage"
  )
)
```

For more control — including a tooltip — use `xl_cell_general()` directly.
Supply a character `value` for the display text and a named-list `hyperlink`
for the URL and tooltip:

```{r hyperlink-detailed}
df4 <- data.frame(x = 1:2)
df4$link <- c(
  xl_cell_general(
    value     = "UCLA website",
    hyperlink = list(url = "http://www.ucla.edu", tooltip = "Opens UCLA homepage")
  ),
  xl_cell_general(hyperlink = "http://www.berkeley.edu")
)
```

`NA` URLs produce blank cells with no hyperlink:

```{r hyperlink-na}
urls <- c("http://www.ucla.edu", NA, "http://www.caltech.edu")
df5 <- data.frame(name = c("UCLA", "???", "Caltech"))
df5$url <- xl_hyperlink(urls)
```

## 3. Value-only cells

`xl_cell_general()` with only a `value` argument writes a plain cell of
whatever R type is provided. This is useful when a column contains a single
scalar that should be recycled:

```{r value-only}
df6 <- data.frame(x = 1:4)
df6$constant <- xl_cell_general(value = 99L)  # integer 99 in every row
```

Supported value types include `numeric`, `integer`, `logical`, `character`,
`Date`, and `POSIXct`.

## 4. Mixed-type columns

Because each element of an `xl_cell_general` object represents one cell, you
can combine different types in a single column by concatenating individual
cells with `c()`:

```{r mixed-type}
mixed <- c(
  xl_cell_general(value = 1.5),                  # numeric
  xl_cell_general(value = "see note"),            # string
  xl_cell_general(formula = "=A1+A2"),            # formula
  xl_cell_general(hyperlink = "https://cran.r-project.org")  # hyperlink
)

df7 <- data.frame(row = 1:4)
df7$mixed_col <- mixed
```

A list passed to `value` allows different R types per cell in one call:

```{r mixed-value-list}
df8 <- data.frame(x = 1:3)
df8$v <- xl_cell_general(value = list(42L, "hello", TRUE))
```

## 5. Backward compatibility

All existing `xl_formula()` and `xl_hyperlink()` call sites continue to work
unchanged. Both functions now return an `xl_cell_general` object instead of a
plain character vector, but the writing behaviour is identical:

```{r compat}
# These calls are identical to writexl < 1.5.5
df9 <- data.frame(
  name    = c("UCLA", "Berkeley"),
  founded = c(1919, 1868),
  website = xl_hyperlink(c("http://www.ucla.edu", "http://www.berkeley.edu"),
                          "homepage")
)
df9$age <- xl_formula('=(YEAR(TODAY()) - INDIRECT("B" & ROW()))')

path <- write_xlsx(df9)
file.exists(path)
```
