---
title: "Writing Special Cell Types with writexl"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Writing Special Cell Types with writexl}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(writexl)
```

The `writexl` package exports plain data frame values (numbers, strings,
booleans, dates) automatically. For anything more specialised — formulas,
hyperlinks, or mixed-type columns — the package provides the `xl_cell_general`
family of helpers.

## 1. Formulas

Use `xl_formula()` to mark character strings as Excel formulas. Each formula
**must** start with `"="`. A length-1 formula is automatically recycled to fill
the whole column.

```{r formulas}
df <- data.frame(
  name    = c("UCLA", "Berkeley", "Caltech"),
  founded = c(1919, 1868, 1891)
)
df$age <- xl_formula('=(YEAR(TODAY()) - INDIRECT("B" & ROW()))')
```

You can also supply a **pre-calculated result** so that xlsx readers that do
not recalculate on open still show a sensible value:

```{r formula-precalc}
cells <- c(
  xl_cell_general(value = 55.0, formula = "=SUM(B2:B4)"),
  xl_cell_general(value = 10.5, formula = "=AVERAGE(B2:B4)")
)
df2 <- data.frame(label = c("Sum", "Mean"))
df2$result <- cells
```

When `value` is numeric the package uses `worksheet_write_formula_num()` so
the stored value is visible even without recalculation. When `value` is a
character string, `worksheet_write_formula_str()` is used instead.

## 2. Hyperlinks

Use `xl_hyperlink()` for clickable URLs. Optionally supply a `name` vector
for the display text shown in the cell:

```{r hyperlinks}
df3 <- data.frame(
  name = c("UCLA", "Berkeley"),
  url  = xl_hyperlink(
    c("http://www.ucla.edu", "http://www.berkeley.edu"),
    name = "homepage"
  )
)
```

For more control — including a tooltip — use `xl_cell_general()` directly.
Supply a character `value` for the display text and a named-list `hyperlink`
for the URL and tooltip:

```{r hyperlink-detailed}
df4 <- data.frame(x = 1:2)
df4$link <- c(
  xl_cell_general(
    value     = "UCLA website",
    hyperlink = list(url = "http://www.ucla.edu", tooltip = "Opens UCLA homepage")
  ),
  xl_cell_general(hyperlink = "http://www.berkeley.edu")
)
```

`NA` URLs produce blank cells with no hyperlink:

```{r hyperlink-na}
urls <- c("http://www.ucla.edu", NA, "http://www.caltech.edu")
df5 <- data.frame(name = c("UCLA", "???", "Caltech"))
df5$url <- xl_hyperlink(urls)
```

## 3. Value-only cells

`xl_cell_general()` with only a `value` argument writes a plain cell of
whatever R type is provided. This is useful when a column contains a single
scalar that should be recycled:

```{r value-only}
df6 <- data.frame(x = 1:4)
df6$constant <- xl_cell_general(value = 99L)  # integer 99 in every row
```

Supported value types include `numeric`, `integer`, `logical`, `character`,
`Date`, and `POSIXct`.

## 4. Mixed-type columns

Because each element of an `xl_cell_general` object represents one cell, you
can combine different types in a single column by concatenating individual
cells with `c()`:

```{r mixed-type}
mixed <- c(
  xl_cell_general(value = 1.5),                  # numeric
  xl_cell_general(value = "see note"),            # string
  xl_cell_general(formula = "=A1+A2"),            # formula
  xl_cell_general(hyperlink = "https://cran.r-project.org")  # hyperlink
)

df7 <- data.frame(row = 1:4)
df7$mixed_col <- mixed
```

A list passed to `value` allows different R types per cell in one call:

```{r mixed-value-list}
df8 <- data.frame(x = 1:3)
df8$v <- xl_cell_general(value = list(42L, "hello", TRUE))
```

## 5. Formatted empty cells

A cell whose value is `NA` normally writes nothing at all, because Excel
ignores blank cells that carry no formatting. Give one a `format`, though, and
the cell is written as a *formatted blank* so the styling survives — useful for
shading gaps in a table:

```{r blank-cells}
df_blank <- data.frame(quarter = c("Q1", "Q2", "Q3"))
df_blank$actual <- xl_cell_general(
  value  = list(101.5, NA, 98.2),
  format = xl_fill(background = "#FFF2CC")
)
```

The middle cell is empty but still shaded. This applies to `xl_cell_general()`
cells; `NA` in a plain data frame column is left to the column's own format,
which Excel already applies to empty cells.

## 6. Array and dynamic array formulas

`array` and `dynamic` control how a formula is stored. A **dynamic** formula is
the modern spelling: you write it to one cell and Excel spills the result over
as many cells as it needs.

```{r dynamic-formula}
df_dyn <- data.frame(city = c("Perth", "Hobart", "Perth", "Darwin"))
df_dyn$unique_cities <- xl_cell_general(
  formula = c("=UNIQUE(A2:A5)", NA, NA, NA),
  dynamic = TRUE
)
```

An **array** formula is the legacy Ctrl-Shift-Enter form, most often used for a
single-cell aggregate:

```{r array-formula}
df_arr <- data.frame(qty = c(2, 5, 3), price = c(10, 4, 7))
df_arr$total <- xl_cell_general(
  formula = c("=SUM(A2:A4*B2:B4)", NA, NA),
  array   = TRUE
)
```

Either may carry a pre-calculated numeric result via `value`, so the workbook
displays a value without Excel recalculating on open. A *character* result is
not possible: Excel stores no cached string result for an array formula, and
supplying one is an error rather than a silent downgrade.

The flags are inert on a cell that has no formula, so a single `array = TRUE`
recycles across a column that mixes formula and value cells.

`array_range` declares the extent of a legacy array formula in the rare case it
must be stated. It has to start at the cell holding the formula and extend into
cells the sheet does not otherwise write, because the range is padded on write
and an overlap would have the padding and your data overwrite each other.
Supplying it also forces the workbook out of its memory-efficient row-streaming
mode. Prefer a single-cell `dynamic` formula wherever you can.

## 7. Rich strings: several fonts in one cell

`xl_rich_string()` splits a single cell's text into runs, each drawn in its own
font. Pass a bare string for an unformatted run, or `xl_rich_run()` to give one
a font:

```{r rich-string}
df_rich <- data.frame(id = 1L)
df_rich$label <- xl_cell_general(
  value = xl_rich_string(
    "This is ",
    xl_rich_run("bold", xl_font(bold = TRUE)),
    " and this is ",
    xl_rich_run("red italic", xl_font(color = "red", italic = TRUE))
  )
)
```

A run is drawn with a **font only**, so only the `xl_font()` properties apply;
supplying a fill, border, alignment or number format on a run warns and is
ignored. A format on the *cell* (`xl_cell_general(format =)`) still works
normally and applies to the whole cell.

Excel needs at least two runs — a string with one format is an ordinary
character value, so pass it as one.

## 8. Backward compatibility

All existing `xl_formula()` and `xl_hyperlink()` call sites continue to work
unchanged. Both functions now return an `xl_cell_general` object instead of a
plain character vector, but the writing behaviour is identical:

```{r compat}
# These calls are identical to writexl < 1.5.5
df9 <- data.frame(
  name    = c("UCLA", "Berkeley"),
  founded = c(1919, 1868),
  website = xl_hyperlink(c("http://www.ucla.edu", "http://www.berkeley.edu"),
                          "homepage")
)
df9$age <- xl_formula('=(YEAR(TODAY()) - INDIRECT("B" & ROW()))')

path <- write_xlsx(df9)
file.exists(path)
```
