---
title: "Charts and images"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Charts and images}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(writexl)
sales <- data.frame(quarter = c("Q1", "Q2", "Q3", "Q4"),
                    revenue = c(120, 145, 133, 180),
                    cost    = c(90, 110, 105, 130))
```

## A chart

`xl_chart()` puts a chart on a sheet; `xl_chart_series()` says what it plots.

```{r}
sheet <- xl_sheet(sales,
                  chart = xl_chart("column",
                                   xl_chart_series(values = list(cols = "revenue")),
                                   title = "Revenue by quarter", at = "E2"))
path <- write_xlsx(list(Sales = sheet), tempfile(fileext = ".xlsx"))
```

All 22 types Excel offers are available: `"column"`, `"bar"`, `"line"`,
`"pie"`, `"doughnut"`, `"area"`, `"scatter"`, `"radar"`, and the stacked,
percent-stacked, smoothed and marker variants.

### Naming a range

A series range is written either **by column**, `list(cols = "revenue")`, or as
**A1 text**, `"Data!B2:B5"`. Prefer the column form: it is checked against the
data frame, moves with the column rather than with a letter, and accounts for
the header row itself. `rows` narrows it, and `sheet` points at another sheet:

```{r}
xl_chart_series(values = list(sheet = "Sales", cols = "revenue", rows = 1:2))
```

A range reaching past the data is an error rather than a chart with blanks on
the end --- a chart of the wrong numbers looks plausible, which is the worst
way to be wrong.

A series that plots a column is **named after that column's header**, which is
what Excel does when you chart a column along with its header. The header is
referenced, not copied, so renaming it in Excel renames the series.
`name = FALSE` leaves it unnamed, and a string is always taken literally --- a
series may legitimately be called `"Q1!"` --- so to take a name from a cell,
give a spec: `list(header = "cost")`, or `list(rows = 1, cols = 1)`.

### Styling

Charts reuse the format objects from [Formatting cells](b-formatting.html).
`xl_border()` becomes the line, `xl_fill()` the fill or pattern, `xl_font()`
the text:

```{r}
xl_chart("line",
         xl_chart_series(values = list(cols = "revenue"),
                         format = xl_border(all = "dashed", color = "#4472C4")),
         title = "Revenue", title_format = xl_font(size = 14, bold = TRUE))
```

Each part takes the groups it can use and **refuses the rest by name**. A
series is a shape, so it takes a line and a fill but no font; a title is text,
so it takes a font and nothing else. `xl_fill()` and `xl_border()` also take
`transparency`, which charts honour and cells ignore, and chart lines have
fewer dash styles than cell borders because a chart line has no width. None of
this is dropped quietly.

`style` picks one of Excel's 48 built-in chart styles.

## Axes

`xl_chart_axis()` covers everything an axis has --- the title and its font,
tick labels with their number format, bounds, a log scale, tick marks and their
spacing, gridlines, and the axis line itself:

```{r}
xl_chart("column", xl_chart_series(values = list(cols = "revenue")),
         x_axis = xl_chart_axis(name = "Quarter", major_tick = "none"),
         y_axis = xl_chart_axis(name = "Revenue", min = 0,
                                num_format = "$#,##0",
                                major_gridlines_format =
                                  xl_border(all = "dotted", color = "gray")))
```

There are two kinds of axis and several options belong to one of them. A
**scatter** chart plots numbers against numbers, so both its axes are value
axes; every other type has a **category** x axis and a value y axis --- bar
charts included, since Excel draws their categories up the side but does not
rename the axes.

| value axes | category axes |
|---|---|
| `min`, `max`, `log_base`, `major_unit`, `minor_unit`, `display_units`, `display_units_visible` | `position`, `label_align`, `interval_unit`, `interval_tick` |

Setting a minimum on a category axis is accepted by Excel and then ignored, so
writexl refuses it and names the axis that would work. Pie and doughnut charts
have no axes at all.

## The parts of a series

A series can carry a marker at each point, labels beside them, a trendline
through them and error bars on them, and individual points can be styled:

```{r}
xl_chart_series(
  values       = list(cols = "revenue"),
  marker       = xl_chart_marker(type = "circle", size = 7),
  labels       = xl_chart_labels(num_format = "$#,##0", position = "above"),
  trendline    = xl_chart_trendline("linear", equation = TRUE),
  y_error_bars = xl_chart_error_bars("percentage", 5))
```

`xl_chart_labels()` with no arguments prints each point's value, which is
Excel's default. Naming any part means the label holds **exactly** those, so
`show_percentage = TRUE` alone gives a percentage with no value beside it.
`custom =` takes a list of `xl_chart_label()`s, one per point, so a single
label can be renamed or removed with `hide = TRUE`. `points =` takes a list of
formats the same way --- one red slice, the rest untouched:

```{r}
xl_chart("pie", xl_chart_series(values = list(cols = "revenue"),
                                points = list(xl_fill(background = "red"),
                                              NULL, NULL, NULL)))
```

Four more of Excel's silent rules are checked here: which **label positions**
each chart type allows, that a **moving average** has no forecast, equation or
R-squared, that an **intercept** applies only to exponential, linear and
polynomial fits, and that an **automatic marker** takes neither a size nor a
format.

## The chart as a whole

```{r}
xl_chart("column", xl_chart_series(values = list(cols = "revenue")),
         legend     = xl_chart_legend(position = "bottom"),
         data_table = xl_chart_table(show_keys = TRUE),
         plot_area_format  = xl_fill(background = "#F8F8F8"),
         chart_area_format = xl_border(all = "thin", color = "gray"))
```

`xl_chart_legend(position = "none")` removes the legend, and `delete_series`
leaves one series out of the key while still plotting it. Anything placed by
hand takes a `layout` rather than an `at` --- `title_layout`,
`plot_area_layout`, `xl_chart_axis(name_layout =)` and
`xl_chart_legend(layout =)` are fractions of the chart, where `at` means a cell
everywhere else. `show_blanks` decides what an empty cell does, and
`show_hidden_data` plots rows Excel would leave out.

Six options belong to one family of chart: the doughnut `hole_size`, pie
`rotation`, `drop_lines`, `high_low_lines`, `up_down_bars`, and the bar
`series_gap` and `series_overlap`. Each takes `TRUE` for Excel's own drawing or
a format to style it, and each is refused on a type that has no such feature.

A scatter series must have `categories` --- they are its x axis, and
libxlsxwriter crashes without them.

## A chart on a sheet of its own

`xl_chartsheet()` is a tab holding one chart and no cells:

```{r}
chart <- xl_chart("column",
                  xl_chart_series(values = list(sheet = "Data",
                                                cols = "revenue"),
                                  categories = list(sheet = "Data",
                                                    cols = "quarter")),
                  title = "Revenue")
path <- write_xlsx(list(Data = sales, Overview = xl_chartsheet(chart)),
                   tempfile(fileext = ".xlsx"))
```

Every range must name its sheet, since a chartsheet has no cells for
`list(cols = )` to resolve against. It supports a fraction of what a worksheet
does --- of `xl_page_setup()` the orientation, paper, margins, header and
footer; of `xl_sheet_view()` the four tab-state options --- plus `tab_color`,
`zoom` and `protect`. The rest is refused rather than dropped.

## Images

`xl_image()` places a picture, anchored to a cell:

```{r eval = isTRUE(capabilities("png"))}
card <- as.raster(matrix(c("red", "blue", "green", "gold"), nrow = 2))
xl_sheet(sales, image = xl_image(card, at = "G2", scale = 20))
```

The image may be a **file path**, a **raw vector** of encoded bytes, or an
**in-memory picture** --- a `raster`, a colour matrix, an RGB array or a
`nativeRaster`. Anything `rasterImage()` can draw can be written, so a plot
never has to touch the disk; that path uses R's PNG device, which is why
`grDevices` is a suggested package.

PNG, JPEG, GIF and BMP are supported, and the format is read from the file's
own bytes rather than its extension, so a mislabelled `.png` is caught rather
than misfiled. `scale`, `offset` and `position` control size and placement,
`description` and `decorative` set the alt text, and `url` makes the image a
hyperlink.

An image can also go in the printed header or footer, matched to a `&G`
placeholder, or be tiled behind the cells as a screen watermark:

```{r eval = FALSE}
xl_page_setup(header = "&L&G", header_image = list(left = logo))
xl_sheet(sales, background = logo)
```

`embed = TRUE` puts the image *inside* a cell, the Excel 365 "place in cell"
feature; older Excel shows `#VALUE!` in its place.

Two combinations are refused, because libxlsxwriter miscounts them and Excel
repairs the file: an embedded image alongside any other image, and a
header/footer or background image on a sheet *before* one with a floating image
or a chart. Both errors name the sheets and the arrangement that works.
