---
title: "Formatting and workbook properties"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formatting and workbook properties}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(writexl)
```

`writexl` writes plain data frames with sensible defaults, but you can also
control formatting at three levels that share one vocabulary:

* **cell** formatting, attached to columns with `xl_cell_general()`;
* **worksheet** layout, attached to a sheet with `xl_sheet()`;
* **workbook** defaults and metadata, attached with `xl_workbook()` /
  `xl_properties()`.

All three reuse the same [format engine](#the-format-engine), so once you know
how to build an `xl_format`, you can apply it anywhere.

## The format engine

A format is built from **group constructors**, each covering one family of
Excel properties. Every argument defaults to "unset", so you only specify what
you want to change.

```{r}
xl_font(bold = TRUE, color = "navy", size = 12)
xl_fill(background = "#FFF2CC")
xl_border(all = "thin", color = "gray")
xl_align(horizontal = "center", vertical = "top", wrap = TRUE)
xl_num_format("#,##0.00")
xl_protection(locked = FALSE)
```

Colors accept R color names (`"navy"`), hex strings (`"#FF0000"`), or integers,
via `xl_color()`.

Each group constructor already returns a complete `xl_format`, so a single
group can be used on its own. Combine groups with `+` (or `xl_format()`), which
merges **property by property** — later values win, but nothing you did not
touch is lost:

```{r}
fmt <- xl_font(bold = TRUE) + xl_fill(background = "yellow") + xl_num_format("0.0%")
fmt
```

## Cell formatting

Attach a format to a column of values with `xl_cell_general()`. Pass one format
for the whole column, or a list of formats for individual cells.

```{r}
df <- data.frame(item = c("Widget", "Gadget", "Gizmo"))
df$price <- xl_cell_general(
  value  = c(1234.5, 67.25, 890),
  format = xl_num_format("$#,##0.00") + xl_font(color = "darkgreen")
)
df$flag <- xl_cell_general(
  value  = c(10, -5, 3),
  format = list(
    xl_fill(background = "lightgreen"),
    xl_fill(background = "salmon"),
    xl_fill(background = "lightgreen")
  )
)
path <- write_xlsx(df, tempfile(fileext = ".xlsx"))
```

`xl_formula()` and `xl_hyperlink_cell()` also take a `format` argument. When a
cell holds a date or time and its format sets no number format, the default
date/time format is applied automatically.

## Worksheet layout

`xl_sheet()` wraps a data frame with column and row specifications and
sheet-level options. `xl_col_spec()` and `xl_row_spec()` are themselves
`xl_format` objects (so they combine with `+`) that additionally carry a target
and geometry.

```{r}
sheet <- xl_sheet(
  data.frame(
    date    = as.Date("2024-01-01") + 0:2,
    revenue = c(1000.5, 2000.25, 1500.75)
  ),
  cols = list(
    xl_col_spec("date", width = 12),
    xl_col_spec("revenue", width = 14,
                format = xl_num_format("#,##0.00") + xl_font(bold = TRUE))
  ),
  rows      = xl_row_spec(1, height = 20),
  freeze    = "A2",          # freeze the header row
  tab_color = "steelblue",
  zoom      = 120
)
path <- write_xlsx(list(Sales = sheet), tempfile(fileext = ".xlsx"))
```

`xl_sheet()` also adds **autofilters** and **worksheet protection**:

```{r}
locked <- xl_sheet(
  data.frame(id = 1:3, note = c("a", "b", "c")),
  autofilter = TRUE,                       # or a range like "A1:B4"
  protect    = list(password = "secret",   # protect, but still allow sorting
                    sort = TRUE)
)
path <- write_xlsx(list(Locked = locked), tempfile(fileext = ".xlsx"))
```

Cell locking set with `xl_protection()` only takes effect once the sheet is
protected.

Set `auto_colwidth = TRUE` to size every column to its contents (a
character-count estimate, since the xlsx format has no true "AutoFit");
columns given an explicit width via `xl_col_spec()` are left as-is.

Geometry can be given in **pixels** instead of Excel's character units and
points, which is easier to reason about when sizing a column around an image
or a fixed layout. Give one unit or the other, not both:

```{r}
xl_col_spec("thumbnail", width_pixels = 100)
xl_row_spec(1, height_pixels = 40)
```

Excel stores character units and points, so the value is converted on the way
in — a column set to 100 pixels reads back as 13.57 character units, and
renders at 100 pixels again.

### Grouping and outlines

`level` in `xl_col_spec()` / `xl_row_spec()` puts columns or rows into an
outline group, which Excel draws with the +/- controls that collapse it.
`xl_outline()` changes how those controls are *drawn* — never which rows are
grouped — and its defaults already match Excel's, so it is rarely needed:

```{r}
grouped <- xl_sheet(
  data.frame(q1 = 1:2, q2 = 3:4, total = 4:5),
  cols    = list(xl_col_spec(c("q1", "q2"), level = 1)),
  outline = xl_outline(symbols_right = FALSE)   # summary column on the left
)
path <- write_xlsx(list(Grouped = grouped), tempfile(fileext = ".xlsx"))
```

### Silencing Excel's error indicators

Excel puts a green triangle in cells it thinks are wrong — most often a number
deliberately stored as text. `ignore_errors` turns that off for a range, one
entry per error type:

```{r}
codes <- xl_sheet(
  data.frame(zip = c("01234", "02138")),
  ignore_errors = list(number_stored_as_text = "A2:A3")
)
path <- write_xlsx(list(Codes = codes), tempfile(fileext = ".xlsx"))
```

The range takes any spelling the other range arguments accept, so
`list(cols = "zip")` would cover the whole column. The available types are
`number_stored_as_text`, `eval_error`, `formula_differs`, `formula_range`,
`formula_unlocked`, `empty_cell_reference`, `list_data_validation`,
`calculated_column` and `two_digit_text_year`.

Columns can be targeted by name or position. A plain data frame passed
alongside an `xl_sheet` behaves exactly as before:

```{r}
path <- write_xlsx(list(
  Styled = sheet,
  Raw    = data.frame(a = 1:3, b = letters[1:3])
), tempfile(fileext = ".xlsx"))
```

## Workbook defaults and metadata

`xl_workbook()` binds sheets to `xl_properties()`, which holds document
metadata plus the formatting defaults that would otherwise be built in. Because
those defaults are ordinary `xl_format` objects, you can change them — for
example, restyle the header row or set a workbook-wide default font.

```{r}
wb <- xl_workbook(
  list(Report = data.frame(quarter = c("Q1", "Q2"), sales = c(12000, 15000))),
  properties = xl_properties(
    title  = "Quarterly Report",
    author = "Finance Team",
    company = "Example Corp",
    custom = list(Project = "Alpha", Reviewed = TRUE),
    default_format = xl_font(name = "Calibri"),
    header_format  = xl_font(bold = TRUE, color = "white") +
                     xl_fill(background = "navy"),
    read_only = TRUE
  )
)
path <- write_xlsx(wb, tempfile(fileext = ".xlsx"))
```

Formats cascade from the workbook down: the effective format for a cell is
`default_format` merged with the column/row/header/hyperlink default, then with
the cell's own format. Identical resulting formats are written once, so
applying the same format to thousands of cells stays compact.

Note that a workbook-wide `default_format` is *emulated* (it is applied beneath
every cell) because the underlying library has no native "Normal style"
setter; it does not change Excel's built-in Normal style.

## Comments

Cell comments (notes) are attached through `xl_cell_general()`'s `comment`
argument. The simplest form is just text — a character vector, one comment per
cell (`NA` for none):

```{r}
df <- data.frame(item = c("Widget", "Gadget", "Gizmo"))
df$item <- xl_cell_general(
  value   = c("Widget", "Gadget", "Gizmo"),
  comment = c("Discontinued", NA, "Low stock")
)
path <- write_xlsx(df, tempfile(fileext = ".xlsx"))
```

`xl_comment()` adds options — an author, initial visibility, and box
size/position — and styling. Excel comment boxes support only a background
color and a font name/size/family, so styling is supplied by reusing the format
engine via `format`; any other format property triggers a warning.

```{r}
note <- xl_comment(
  "Please verify with Finance",
  author  = "Reviewer",
  visible = TRUE,
  format  = xl_font(name = "Arial", size = 10) + xl_fill(background = "lightyellow")
)
d <- data.frame(x = 1)
d$x <- xl_cell_general(value = 1234.5, comment = note)
path <- write_xlsx(d, tempfile(fileext = ".xlsx"))
```

A default author, and showing all of a sheet's comments at once, are set on
`xl_sheet()` (comment defaults are worksheet-scoped in Excel). A comment's own
`author` overrides the sheet default, and `visible` overrides `show_comments`:

```{r}
sheet <- xl_sheet(df, comment_author = "QA", show_comments = TRUE)
path <- write_xlsx(list(Data = sheet), tempfile(fileext = ".xlsx"))
```

## Page setup and printing

`xl_page_setup()` collects everything on Excel's "Page Layout" ribbon. Pass it
as `xl_sheet(page = )`. None of it touches the cell data --- it changes only
how the sheet prints and how it looks in page-break preview.

```{r}
sheet <- xl_sheet(df, page = xl_page_setup(
  orientation = "landscape",
  paper       = "A4",
  margins     = c(left = 0.5, right = 0.5),
  fit_to      = c(width = 1, height = 0),   # one page wide, any number tall
  header      = "&LQuarterly report&RPage &P of &N"
))
path <- write_xlsx(list(Data = sheet), tempfile(fileext = ".xlsx"))
```

`margins` takes one value for all four sides, four values in the order left,
right, top, bottom, or a named vector naming only the sides you care about.
`paper` takes a friendly name or an Excel paper-type integer for the envelope
and specialist sizes.

`fit_to` scales the printout to a number of pages, where `0` means "as many as
needed in that direction". `scale` sets a plain percentage instead; Excel
ignores it when `fit_to` is set.

### Headers and footers

Header and footer text uses Excel's own codes: `&L`, `&C` and `&R` start the
left, centre and right sections, `&P` is the page number, `&N` the page count,
`&D` the date, `&A` the sheet name, and `&&` a literal ampersand. The whole
string is limited to 255 characters.

```{r}
xl_page_setup(header = "&LDraft&CQ1&R&D", footer = "&CPage &P of &N")
```

Header and footer *images* are not supported yet, so an `&G` or `&[Picture]`
placeholder is rejected.

### Repeating headings and choosing what prints

```{r}
sheet <- xl_sheet(df, page = xl_page_setup(
  print_area  = "A1:D50",
  repeat_rows = 1,          # repeat the header row on every page
  repeat_cols = "A:A"       # and the first column
))
```

`print_area` takes any range writexl accepts, including a
`list(rows = , cols = )` spec. `repeat_rows` and `repeat_cols` take either a
count or a range string.

Note that these count **sheet** rows from 1, with the header row included ---
unlike `xl_row_spec()`, which indexes data rows. Repeating the header row is
the usual reason to use them, so it has to be nameable.

### Manual page breaks

`h_breaks` and `v_breaks` give 1-based positions at which a new page starts, so
`h_breaks = 21` breaks between rows 20 and 21.

```{r}
xl_page_setup(h_breaks = c(21, 41), v_breaks = "D")
```

Excel ignores manual breaks when `fit_to` is set, and writexl warns if you ask
for both.

## Sheet tabs and the opening view

`xl_sheet()` also controls the tab strip and what the user sees when the sheet
opens.

```{r}
wb <- list(
  Summary = xl_sheet(df, view = xl_sheet_view(active = TRUE, selection = "B2")),
  Detail  = xl_sheet(df, view = xl_sheet_view(hide_zero = TRUE)),
  Working = xl_sheet(df, view = xl_sheet_view(visible = FALSE))     # hidden tab
)
path <- write_xlsx(wb, tempfile(fileext = ".xlsx"))
```

`active` picks the tab Excel opens on, `selected` adds a tab to the selected
group, `visible = FALSE` hides one, and `first_tab` chooses the leftmost tab in
the strip.

These are the only worksheet settings whose rules span the whole workbook, and
writexl checks them before writing anything:

* a hidden sheet cannot be active or selected;
* at most one sheet may be active;
* the first sheet cannot be hidden unless another is made active, because Excel
  opens on the first sheet by default;
* at least one sheet must stay visible, or Excel refuses to open the file.

Each failure names the sheet at fault. libxlsxwriter enforces none of them, so
without these checks a bad combination would produce a workbook that simply
does not open.

### Where the sheet is scrolled to

`selection` sets the selected cell or range, and `top_left` the cell scrolled to
the top-left of the window:

```{r}
xl_sheet(df, view = xl_sheet_view(selection = "B2:D10", top_left = "A2"))
```

### Freezing and splitting

`freeze` locks rows and columns in place; `split` gives a movable divider with
independent scroll bars instead. They are mutually exclusive.

```{r}
xl_sheet(df, freeze = "A2")   # keep the header row visible
xl_sheet(df, view = xl_sheet_view(split  = "B3"))   # split above row 3 and left of column B
```

Note that libxlsxwriter positions a split by *distance*, in row-height and
column-width units, rather than by row and column — `15` means one default-height
row, `8.43` one default-width column. writexl converts the cell reference for
you using the sheet's actual row heights and column widths, so a split still
lands where you asked on a sheet whose rows or columns have been resized. Pass
`list(vertical = , horizontal = )` if you would rather give the units directly.

## Merged cells

`xl_merge()` merges a rectangle of cells into one, as Excel's "Merge and
Centre" does. Pass one or a list of them as `xl_sheet(merge = )`.

```{r}
sheet <- xl_sheet(df, merge = xl_merge(
  "A5:B5", "Total",
  format = xl_align(horizontal = "center") + xl_font(bold = TRUE)
))
path <- write_xlsx(list(Data = sheet), tempfile(fileext = ".xlsx"))
```

A merged range holds a single value, so `xl_merge()` carries its own `text`
rather than taking it from the data frame. Merges are applied after the sheet's
rows are written, so merging over cells the data frame filled keeps only the
merged text — exactly as merging in Excel discards everything but the top-left
value.

Excel has no single-cell merge, so a range covering one cell is an error.

Because a merge writes back over rows already emitted, any merge turns off the
memory-efficient row-streaming mode for the whole workbook. That is invisible
apart from a slightly different internal string storage, but it does mean the
whole sheet is held in memory while writing.

## Data validation

`xl_validation()` restricts what may be typed into a range, and pairs with
`xl_sheet(validation = )`.

```{r}
sheet <- xl_sheet(df, validation = list(
  xl_validation("A2:A4", type = "integer", min = 1, max = 10,
                error_message = "Enter a whole number from 1 to 10"),
  xl_validation("B2:B4", list = c("open", "high", "close"))
))
path <- write_xlsx(list(Data = sheet), tempfile(fileext = ".xlsx"))
```

Excel has seventeen internal validation types; writexl exposes five kinds ---
`"integer"`, `"decimal"`, `"date"`, `"time"` and `"length"` --- plus `list`,
`"custom"` and `"any"`, and works out the rest from what you pass. A limit that
is a string beginning `"="` is a formula, and a `Date` or `POSIXct` is a
date/time bound:

```{r}
xl_validation("C2:C4", type = "integer", criteria = ">", value = "=$Z$1")
xl_validation("D2:D4", type = "date", criteria = ">=",
              value = as.Date("2024-01-01"))
```

`criteria` is required for the five comparison kinds and must be omitted for
`list`, `"custom"` and `"any"`, which carry their own meaning. Giving both
`min` and `max` implies `"between"`, so the criteria can be left out there too.

`input_title` / `input_message` show a tooltip when the cell is selected, and
`error_title` / `error_message` appear on a bad entry; `error_type` chooses
whether Excel refuses the entry (`"stop"`, the default) or merely warns. Excel
limits titles to 32 characters and messages to 255, and a dropdown's choices to
255 characters once joined by commas --- put longer lists in a range and pass a
`"=..."` formula instead.

## Conditional formatting

Four constructors cover Excel's conditional formatting, and all of them go into
`xl_sheet(conditional = )`:

```{r}
sheet <- xl_sheet(df, conditional = list(
  xl_cond_cell("A2:A4", criteria = ">", value = 2,
               format = xl_fill(background = "#FFC7CE")),
  xl_cond_scale("B2:B4", colors = c("red", "yellow", "green"))
))
path <- write_xlsx(list(Data = sheet), tempfile(fileext = ".xlsx"))
```

`xl_cond_cell()` pairs a condition with a format. The condition can be a
comparison, a text match, a time period, above/below average, top/bottom N,
duplicates, uniques, blanks, errors, or an arbitrary formula:

```{r}
xl_cond_cell("A2:A100", criteria = "contains", value = "urgent")
xl_cond_cell("A2:A100", type = "duplicate", format = xl_font(bold = TRUE))
xl_cond_cell("A2:A100", type = "formula", value = "=$B2>$C2")
```

Each rule type accepts its own set of criteria — text criteria only make sense
for a text rule, time-period criteria only for a time rule. Pairing them wrongly
is an error, because Excel would accept the file and then silently ignore the
rule. Where the two are redundant the type is inferred, so
`criteria = "contains"` alone is enough.

A rule's format is an ordinary `xl_format`, but Excel stores it as a
*differential* format rather than a cell style, so only the properties that make
sense as an override are used.

### Scales, bars and icons

```{r}
xl_cond_scale("B2:B100", colors = c("white", "steelblue"))  # two-colour
xl_cond_bar("C2:C100", color = "steelblue", solid = TRUE)
xl_cond_icons("D2:D100", style = "3_traffic_lights")
```

`xl_cond_scale()` takes two or three colours, low to high, optionally with the
values and rule types marking where each sits. `xl_cond_bar()` draws in-cell
data bars, with control over the negative and border colours, the axis and the
direction. `xl_cond_icons()` selects one of Excel's seventeen built-in icon
sets — these are drawn by Excel itself, so nothing is embedded in the file.

## Filtering

`xl_sheet(autofilter = TRUE)` adds the filter dropdowns. `xl_filter()` goes
further and sets the criteria on a column:

```{r}
fruit <- data.frame(
  fruit = c("apple", "banana", "cherry", "damson", "elderberry"),
  qty   = c(5, 150, 20, 300, 75),
  stringsAsFactors = FALSE
)
sheet <- xl_sheet(fruit, filter = xl_filter("qty", ">", 100))
path <- write_xlsx(list(Data = sheet), tempfile(fileext = ".xlsx"))
```

That file opens showing only banana and damson.

### Why writexl hides the rows

Excel stores a filter's *criteria* and its *hidden rows* as two independent
things, and it does not apply a filter when a file is opened. A file carrying
criteria alone therefore opens with the funnel icon showing "greater than 100"
and every row still visible — it looks filtered and is not.

So `xl_filter()` does both: it writes the criteria, and it hides the rows the
criteria excludes. That means writexl has to reproduce Excel's own matching
rules, which were established by writing criteria with no hidden rows, opening
the file and pressing Data → Reapply so that Excel computed the match itself.

### Two kinds of filter

The measurements showed that the rules are not a property of the criteria on
its own. A filter is written in one of two forms, and Excel matches them
differently:

| Form | Written for | Excel matches |
|---|---|---|
| **value list** | `"=="` without a wildcard, `"blanks"`, any `list` | the text the cell **displays**, case-insensitively |
| **typed comparison** | everything else, including `"=="` *with* a wildcard | by **type** — numbers to numbers, text to text |

A value list ignores how the value was stored, so all three of these keep the
number `10` *and* the string `"10"`:

```{r}
xl_filter("qty", "==", 10)
xl_filter("qty", "==", "10")
xl_filter("qty", list = "10")
```

A typed comparison does not. On a numeric column `"=="` with `"1*"` keeps
nothing at all, because no number is text — even though `"=="` with `"10"`
keeps the number 10. The wildcard is the only difference, and it changes the
form:

```{r}
xl_filter("qty", "==", "1*")   # typed: matches no number
xl_filter("qty", "==", "10")   # value list: matches the number 10
```

Within a typed comparison, a text cell is never equal to a number — which makes
`"!="` *true* for it. So `xl_filter("fruit", "!=", 10)` keeps every row of a
text column.

The rest is more predictable: text comparison is case-insensitive, `*` and `?`
are wildcards, and **blank** covers an empty cell as well as an empty string,
with `"non-blanks"` its exact complement. Blanks are dropped by every
comparison save one — `"!="` keeps them, since a blank is not equal to
anything.

```{r}
xl_filter("fruit", "==", "ap*")                  # wildcard, case-insensitive
xl_filter("fruit", list = c("apple", "cherry"))  # keep only these
xl_filter("qty", ">", 100, "<", 200)             # two rules, combined with and
xl_filter("fruit", "non-blanks")
```

Filters on different columns combine with AND, as they do in Excel.

### Mixed columns, and what cannot be filtered

A column built with `xl_cell_general()` may hold different types in different
rows, and each cell is matched by its own type:

```{r}
mixed <- data.frame(row = 1:5)
mixed$v <- xl_cell_general(value = list(10, "20", "abc", 30, 20))
```

Here `xl_filter("v", ">", 15)` keeps rows 4 and 5 — the numbers 30 and 20 — but
not the *text* `"20"` in row 2. A value list works the other way: because it
matches displayed text, `xl_filter("v", list = "20")` keeps both row 2 and
row 5.

### Asking without writing

The matching rule is available on its own, so you can see which rows a filter
would leave visible without writing a file:

```{r}
sales <- data.frame(fruit = c("apple", "banana", "cherry"),
                    qty = c(5, 150, 300))
xl_filter_keep(sales, xl_filter("qty", ">", 100))
sales[xl_filter_keep(sales, xl_filter("fruit", "==", "*a*")), ]
```

`xl_sheet(filter =)` uses this same function to decide what to hide, so the two
cannot disagree.

Two things are refused rather than guessed at. A column of **formulas** cannot
be filtered, because writexl does not know what Excel would compute. And a
**date** column cannot be filtered by `"=="` or `list`: those match displayed
text, which depends on the number format, so use a comparison such as `">="`
instead. The same caution applies to a numeric column carrying a custom number
format — writexl compares the General rendering of the number, so a value list
may not match what Excel shows.

## Tables

`xl_table()` turns a range into an Excel table: a named, styled block with
banded rows, a filter dropdown in its header, and structured references in
formulas.

```{r}
sales <- data.frame(fruit = c("apple", "banana", "cherry"),
                    qty   = c(5, 150, 300))
sheet <- xl_sheet(sales, table = xl_table(style = "medium 9"))
path <- write_xlsx(list(Sales = sheet), tempfile(fileext = ".xlsx"))
```

The header captions come from the data frame's column names. That is worth
knowing rather than taking for granted: Excel stores a table's column names
separately from the header cells and refuses to open a file where the two
disagree, so the default is what makes a table safe to add at all. Override one
with `xl_table_column()` and both sides change together.

Styles are a type and a number — `"light 0"` to `"light 21"`, `"medium 1"` to
`"medium 28"`, `"dark 1"` to `"dark 11"` — or `"none"` for an unstyled table.

### Totals and formulas

A total row is added below the data, with each column choosing a function or a
label:

```{r}
totalled <- xl_sheet(sales, table = xl_table(
  total_row = TRUE,
  columns = list(xl_table_column("fruit", total_label = "Total"),
                 xl_table_column("qty", total = "sum"))
))
path <- write_xlsx(list(Sales = totalled), tempfile(fileext = ".xlsx"))
```

The total row is written by Excel's own machinery and the sheet's row plan does
not see it: `auto_colwidth` does not measure it, and an `xl_row_spec()` aimed at
that row will fight it.

A column formula usually uses a structured reference, which names the table —
so give the table a `name` when you write one:

```{r}
projected <- sales
projected$doubled <- NA_real_
sheet <- xl_sheet(projected, table = xl_table(
  name = "Sales",
  columns = xl_table_column("doubled",
                            formula = "=Sales[[#This Row],[qty]]*2")
))
path <- write_xlsx(list(Sales = sheet), tempfile(fileext = ".xlsx"))
```

Left to itself, `xl_table()` generates a name from the sheet name and writes it
explicitly, so it cannot shift when a table is added elsewhere. `name = NA`
leaves the naming to Excel (`Table1`, `Table2`, …), which is fine until a
formula refers to one — writexl warns if you combine the two.

### What a table will not sit alongside

Excel silently repairs each of these, so writexl refuses them instead: a table
overlapping `xl_sheet(autofilter =)` (the table brings its own dropdown), an
`xl_filter()` on a column the table covers, a merged range inside the table, and
two overlapping tables. Each error quotes the ranges involved.

Any table also turns off the memory-efficient row-streaming mode, which
libxlsxwriter refuses to combine with tables at all.

## Images

`xl_image()` places a picture on a sheet, anchored to a cell:

```{r}
logo <- system.file("help", "figures", "logo.png", package = "writexl")
```
```{r eval = nzchar(logo)}
sheet <- xl_sheet(data.frame(product = c("a", "b")),
                  image = xl_image(logo, at = "D2", scale = 0.5))
path <- write_xlsx(list(Catalogue = sheet), tempfile(fileext = ".xlsx"))
```

The image can be a **file path**, a **raw vector** of encoded bytes, or an
**in-memory picture** — a `raster`, a colour matrix, an RGB array, or a
`nativeRaster`. Anything `rasterImage()` can draw can be written, so a plot
never has to touch the disk:

```{r eval = isTRUE(capabilities("png"))}
card <- as.raster(matrix(c("red", "blue", "green", "gold"), nrow = 2))
xl_image(card, at = "B2", scale = 20)
```

PNG, JPEG, GIF and BMP are supported, and the format is read from the file's own
bytes rather than its extension, so a mislabelled `.png` is caught rather than
misfiled. Encoding an in-memory picture uses R's PNG device, which is why
`grDevices` and `graphics` are suggested packages; if they are unavailable the
error tells you how to encode it yourself.

`scale`, `offset` and `position` control size and placement, `description` and
`decorative` set the alt text screen readers use, and `url` makes the image a
hyperlink.

### Headers, footers and backgrounds

An image can also go in the printed header or footer, matched to a `&G`
placeholder, or be tiled behind the cells as a screen watermark:

```{r eval = nzchar(logo)}
xl_page_setup(header = "&L&G&CQuarterly report",
              header_image = list(left = logo))
```

A background is a display-only backdrop — Excel never prints it. Header and
footer images appear only in Page Layout view and Print Preview, not in the
normal grid.

### Embedded images, and what cannot be combined

`embed = TRUE` puts the image *inside* a cell, so it sizes with the cell — the
Excel 365 "place in cell" feature. Older Excel shows `#VALUE!` in its place.

Two combinations are refused, because libxlsxwriter miscounts them and Excel
then repairs the file and drops the image:

* an embedded image in a workbook that holds **any other** image;
* a header/footer or background image on a sheet **before** one with an ordinary
  floating image.

Both errors name the sheets and the arrangement that works. Put the sheets with
floating images first, and give embedded images a workbook of their own.

Only embedded images cost the memory-efficient row-streaming mode; ordinary
floating images leave it on, so a large sheet with a logo stays fast.

## Charts

`xl_chart()` puts a chart on a sheet; `xl_chart_series()` says what it plots.
The chart is anchored to a cell in the same vocabulary an image uses:

```{r}
sales <- data.frame(quarter = c("Q1", "Q2", "Q3", "Q4"),
                    revenue = c(120, 145, 133, 180))
sheet <- xl_sheet(sales,
                  chart = xl_chart("column",
                                   xl_chart_series(values = list(cols = "revenue"),
                                                   categories = list(cols = "quarter"),
                                                   name = "2024"),
                                   title = "Revenue by quarter", at = "D2"))
path <- write_xlsx(list(Sales = sheet), tempfile(fileext = ".xlsx"))
```

All 22 chart types Excel offers are available: `"column"`, `"bar"`, `"line"`,
`"pie"`, `"doughnut"`, `"area"`, `"scatter"`, `"radar"`, and the stacked,
percent-stacked, smoothed and marker variants.

### Naming a range

A series range can be written either way:

* **by column**, `list(cols = "revenue")` — the whole column of the sheet's own
  data, header excluded;
* **as A1 text**, `"B2:B5"` or `"Data!B2:B5"`.

Prefer the column form. It is checked against the data frame, it moves with the
column rather than with a letter, and it accounts for the header row on its own,
so adding a row or reordering the columns does not silently plot the wrong
thing. `rows` narrows it: `list(cols = "revenue", rows = 1:2)` is the first two
data rows.

### Naming a series

A series that plots a column is named after that column's **header cell**,
which is what Excel does when you chart a column along with its header. So the
legend usually needs no arguments at all:

```{r}
xl_chart("column", xl_chart_series(values = list(cols = "revenue")))
```

The header is referenced, not copied, so editing it in Excel renames the
series. `name = FALSE` leaves the series unnamed (Excel then calls it
"Series 1"), and anything else overrides the default.

A series name and a chart title are the exception to the A1 form: a string
there is always taken **literally**, because a series may legitimately be
called `"Q1!"`. To take one from a cell, give a spec. `rows` counts *data*
rows, so the header row has its own spelling:

```{r}
xl_chart("column",
         xl_chart_series(values = list(cols = "revenue"),
                         name = list(header = "cost")),   # another column's
         title = list(header = "quarter"))
```

`list(header = )` names one column's header cell and stands alone; it takes an
optional `sheet` like any other spec. The default applies only where a header
is known to exist — not to an A1 range, which says nothing about whether the
row above it is a header, and not on a workbook written with
`col_names = FALSE`, where naming one is an error rather than a guess.

Either form may name another sheet, which is how a chart gets a sheet to itself:

```{r}
overview <- xl_sheet(data.frame(note = "See the chart"),
                     chart = xl_chart("line",
                                      xl_chart_series(values = list(sheet = "Sales",
                                                                    cols = "revenue"),
                                                      categories = list(sheet = "Sales",
                                                                        cols = "quarter"))))
```

A range that reaches past the data is an error, not a chart with blanks on the
end — a chart of a range that does not hold the numbers you meant looks
plausible, which is the worst way for it to be wrong.

A scatter series must have `categories`: a scatter plots values *against*
categories, so they are its x axis.

### Styling a chart

Charts reuse the format groups from the top of this vignette. `xl_border()`
becomes the line, `xl_fill()` the fill or pattern, and `xl_font()` the text, so
one format object can style a cell and a chart alike:

```{r}
xl_chart("line",
         xl_chart_series(values = list(cols = "revenue"),
                         format = xl_border(all = "dashed", color = "#4472C4") +
                                  xl_fill(background = "#4472C4",
                                          transparency = 40)),
         title = "Revenue", title_format = xl_font(size = 14, bold = TRUE))
```

Each part takes the groups it can use and refuses the rest **by name**. A series
is a shape, so it takes a line and a fill but no font; a title is text, so it
takes a font and nothing else. `xl_fill()` and `xl_border()` take
`transparency`, a percentage charts honour and cells ignore, and chart lines
have fewer dash styles than cell borders because a chart line has no width. A
number format or an alignment is refused outright. None of this is dropped
quietly — a style that cannot be drawn should not be mistaken for one that was.

`title_format` needs a `title`. The title Excel puts on a single-series chart is
generated by Excel and never written to the file, so there would be nothing for
the font to attach to.

`style` selects one of Excel's 48 built-in chart styles.

### Axes

`xl_chart_axis()` describes an axis, and covers everything libxlsxwriter can
set on one — the title, tick labels and their number format, bounds, a log
scale, tick marks and their spacing, gridlines, and the axis line itself. An
axis has four parts that can be styled, so each format argument names the part
it styles — `name_format`, `label_format`, `line_format`, and the two gridline
ones:

```{r}
xl_chart("column", xl_chart_series(values = list(cols = "revenue")),
         x_axis = xl_chart_axis(name = "Quarter", major_tick = "none"),
         y_axis = xl_chart_axis(name = "Revenue", min = 0,
                                num_format = "$#,##0",
                                major_gridlines_format =
                                  xl_border(all = "dotted", color = "gray")))
```

There are two kinds of axis, and several options belong to one of them. A
**scatter** chart plots numbers against numbers, so both of its axes are value
axes; every other type has a **category** x axis and a value y axis — bar
charts included, since Excel draws their categories up the side but does not
rename the axes.

| | value axes | category axes |
|---|---|---|
| | `min`, `max`, `log_base`, `major_unit`, `minor_unit`, `display_units`, `display_units_visible` | `position`, `label_align`, `interval_unit`, `interval_tick` |

Everything else applies to both. Setting a minimum on a category axis is
accepted by Excel and then ignored, so writexl refuses it and names the axis
that would work. Pie and doughnut charts have no axes at all.

`visible = FALSE` hides an axis, `label_position = "none"` keeps the axis but
drops its labels, and styling a gridline switches it on — a styled gridline
that stays hidden is not something anyone means.

### The parts of a series

A series can carry a marker at each point, labels beside them, a trendline
through them, and error bars on them; individual points can be styled one at a
time:

```{r}
xl_chart("line",
         xl_chart_series(
           values = list(cols = "revenue"),
           marker = xl_chart_marker(type = "circle", size = 7),
           labels = xl_chart_labels(num_format = "$#,##0",
                                    position = "above"),
           trendline = xl_chart_trendline("linear", equation = TRUE),
           y_error_bars = xl_chart_error_bars("percentage", 5)))
```

`xl_chart_labels()` with no arguments prints each point's value, which is
Excel's own default. Naming any part means the label holds **exactly** the
parts named, so `show_percentage = TRUE` alone gives a percentage with no value
beside it; `separator` says what joins them when there is more than one. `custom =` takes a list of
`xl_chart_label()`s, one per point, so a single label can be given its own text
or removed with `hide = TRUE`. `points =` takes a list of `xl_format()`s the
same way — one red slice in a pie, the rest left alone:

```{r}
xl_chart("pie", xl_chart_series(values = list(cols = "revenue"),
                                points = list(xl_fill(background = "red"),
                                              NULL, NULL, NULL)))
```

Four more of Excel's silent rules are checked here. Which **label positions**
are available depends on the chart type — `"above"` and `"below"` on line and
scatter charts, `"inside_base"` on bar and column, `"best_fit"` on pie and
doughnut, `"center"` everywhere — and the error lists the ones this chart
allows. A **moving average** has no forecast, equation or R-squared; an
**intercept** applies to exponential, linear and polynomial fits only; and an
**automatic marker** takes neither a size nor a format, since it is Excel's own
default symbol for that series.

### Features that belong to some types only

Excel discards a setting the chart type does not have without saying so, so
writexl checks first and names the types that would work. Only a doughnut has a
`hole_size`; only pie and doughnut `rotation`; up-down bars and high-low lines
are line-only; the series gap and overlap are bar and column only; smoothing is
line and scatter only; and pie and doughnut charts have no axes at all.
