---
title: "Formatting cells"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formatting cells}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(writexl)
```

One kind of object describes appearance everywhere in writexl: an `xl_format`.
The same object styles a cell, a whole column, a chart series or a conditional
rule, so there is one vocabulary to learn rather than five.

## Building a format

Six group constructors, each covering one family of Excel properties. Every
argument defaults to unset, so you name only what you want to change:

```{r}
xl_font(bold = TRUE, color = "navy", size = 12)
xl_fill(background = "#FFF2CC")
xl_border(all = "thin", color = "gray")
xl_align(horizontal = "center", vertical = "top", wrap = TRUE)
xl_num_format("#,##0.00")
xl_protection(locked = FALSE)
```

Colours accept R colour names, hex strings or integers, via `xl_color()`.

Each constructor already returns a complete format, so one can be used alone.
Combine them with `+`, which merges **property by property** --- later values
win, but nothing you did not touch is lost:

```{r}
money <- xl_num_format("$#,##0.00") + xl_font(color = "darkgreen")
money + xl_font(bold = TRUE)     # still green, now bold
```

## Applying it to cells

`xl_cell_general()` attaches a format to values. One format covers the column;
a list gives a format per cell:

```{r}
df <- data.frame(item = c("Widget", "Gadget", "Gizmo"))
df$price <- xl_cell_general(value = c(1234.5, 67.25, 890), format = money)
df$flag  <- xl_cell_general(
  value  = c(10, -5, 3),
  format = list(xl_fill(background = "lightgreen"),
                xl_fill(background = "salmon"),
                xl_fill(background = "lightgreen")))
path <- write_xlsx(df, tempfile(fileext = ".xlsx"))
```

A cell with a format and no value is a formatted blank --- useful for ruling a
sheet out to a fixed size:

```{r}
xl_cell_general(value = NA, format = xl_fill(background = "#EEEEEE"))
```

Dates and times take the workbook's date format automatically when their own
format sets no number format.

## Applying it to columns and rows

More often you want a whole column. `xl_col_spec()` and `xl_row_spec()` *are*
`xl_format` objects that additionally carry a target and some geometry, so they
combine with `+` like any other format:

```{r}
sheet <- xl_sheet(
  data.frame(date = as.Date("2024-01-01") + 0:2,
             revenue = c(1000.5, 2000.25, 1500.75)),
  cols = list(xl_col_spec("date", width = 12),
              xl_col_spec("revenue", width = 14, format = money)),
  rows = xl_row_spec(1, height = 20))
path <- write_xlsx(list(Sales = sheet), tempfile(fileext = ".xlsx"))
```

Columns are named or indexed; rows are counted from 1 as *data* rows, ignoring
the header.

## Where a format comes from

Formats cascade. What Excel finally applies to a cell is the workbook's
`default_format`, merged with the column or row format, merged with the cell's
own. Identical results are written once, so styling thousands of cells stays
compact. See [Worksheets and workbooks](sheets.html) for the workbook end of
that.

## Conditional formatting

A conditional format is a rule plus a format, applied to a range and evaluated
by Excel rather than by you:

```{r}
scores <- data.frame(name = c("a", "b", "c", "d"), score = c(35, 78, 55, 92))
sheet <- xl_sheet(scores, conditional = list(
  xl_cond_cell(list(cols = "score"), "cell", ">=", 80,
               format = xl_fill(background = "lightgreen")),
  xl_cond_cell(list(cols = "score"), "cell", "<", 50,
               format = xl_fill(background = "salmon"))))
path <- write_xlsx(list(Scores = sheet), tempfile(fileext = ".xlsx"))
```

Rules are tried in order, and `stop_if_true` stops at the first match.

The format on a rule is a **differential** format: it says what to change, not
what the cell looks like. A rule setting only a fill leaves the font alone,
which is why these are stored apart from ordinary cell styles.

`type` picks what is being tested --- a cell's value, whether it is in the top
or bottom *n*, above or below average, a duplicate, blank, an error, text
containing something, a date in a relative window, or a formula of your own.

### Scales, bars and icons

Three constructors cover Excel's graphical rules, none of which needs a format:

```{r}
xl_cond_scale(list(cols = "score"), colors = c("salmon", "lightgreen"))
xl_cond_bar(list(cols = "score"), color = "steelblue")
xl_cond_icons(list(cols = "score"), style = "3_traffic_lights")
```

A two-colour scale takes two colours, a three-colour scale three. Data bars
draw behind the value in the cell. Icon sets take one of Excel's built-in
`style`s, optionally reversed or shown without their values.

Every one of these takes its range the same way as anything else in writexl:
`list(cols = )` against the data frame, or A1 text.

## What is elsewhere

* The same format objects style **charts**, translated to what a chart can
  draw: [Charts and images](charts.html).
* **Number formats** are ordinary Excel format strings --- anything the Custom
  box in Excel accepts.
* `xl_protection()` only takes effect once the sheet itself is protected: see
  [Worksheets and workbooks](sheets.html).
