---
title: "Formatting and workbook properties"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formatting and workbook properties}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(writexl)
```

`writexl` writes plain data frames with sensible defaults, but you can also
control formatting at three levels that share one vocabulary:

* **cell** formatting, attached to columns with `xl_cell_general()`;
* **worksheet** layout, attached to a sheet with `xl_sheet()`;
* **workbook** defaults and metadata, attached with `xl_workbook()` /
  `xl_properties()`.

All three reuse the same [format engine](#the-format-engine), so once you know
how to build an `xl_format`, you can apply it anywhere.

## The format engine

A format is built from **group constructors**, each covering one family of
Excel properties. Every argument defaults to "unset", so you only specify what
you want to change.

```{r}
xl_font(bold = TRUE, color = "navy", size = 12)
xl_fill(background = "#FFF2CC")
xl_border(all = "thin", color = "gray")
xl_align(horizontal = "center", vertical = "top", wrap = TRUE)
xl_num_format("#,##0.00")
xl_protection(locked = FALSE)
```

Colors accept R color names (`"navy"`), hex strings (`"#FF0000"`), or integers,
via `xl_color()`.

Each group constructor already returns a complete `xl_format`, so a single
group can be used on its own. Combine groups with `+` (or `xl_format()`), which
merges **property by property** — later values win, but nothing you did not
touch is lost:

```{r}
fmt <- xl_font(bold = TRUE) + xl_fill(background = "yellow") + xl_num_format("0.0%")
fmt
```

## Cell formatting

Attach a format to a column of values with `xl_cell_general()`. Pass one format
for the whole column, or a list of formats for individual cells.

```{r}
df <- data.frame(item = c("Widget", "Gadget", "Gizmo"))
df$price <- xl_cell_general(
  value  = c(1234.5, 67.25, 890),
  format = xl_num_format("$#,##0.00") + xl_font(color = "darkgreen")
)
df$flag <- xl_cell_general(
  value  = c(10, -5, 3),
  format = list(
    xl_fill(background = "lightgreen"),
    xl_fill(background = "salmon"),
    xl_fill(background = "lightgreen")
  )
)
path <- write_xlsx(df, tempfile(fileext = ".xlsx"))
```

`xl_formula()` and `xl_hyperlink_cell()` also take a `format` argument. When a
cell holds a date or time and its format sets no number format, the default
date/time format is applied automatically.

## Worksheet layout

`xl_sheet()` wraps a data frame with column and row specifications and
sheet-level options. `xl_col_spec()` and `xl_row_spec()` are themselves
`xl_format` objects (so they combine with `+`) that additionally carry a target
and geometry.

```{r}
sheet <- xl_sheet(
  data.frame(
    date    = as.Date("2024-01-01") + 0:2,
    revenue = c(1000.5, 2000.25, 1500.75)
  ),
  cols = list(
    xl_col_spec("date", width = 12),
    xl_col_spec("revenue", width = 14,
                format = xl_num_format("#,##0.00") + xl_font(bold = TRUE))
  ),
  rows      = xl_row_spec(1, height = 20),
  freeze    = "A2",          # freeze the header row
  tab_color = "steelblue",
  zoom      = 120
)
path <- write_xlsx(list(Sales = sheet), tempfile(fileext = ".xlsx"))
```

`xl_sheet()` also adds **autofilters** and **worksheet protection**:

```{r}
locked <- xl_sheet(
  data.frame(id = 1:3, note = c("a", "b", "c")),
  autofilter = TRUE,                       # or a range like "A1:B4"
  protect    = list(password = "secret",   # protect, but still allow sorting
                    sort = TRUE)
)
path <- write_xlsx(list(Locked = locked), tempfile(fileext = ".xlsx"))
```

Cell locking set with `xl_protection()` only takes effect once the sheet is
protected.

Set `auto_colwidth = TRUE` to size every column to its contents (a
character-count estimate, since the xlsx format has no true "AutoFit");
columns given an explicit width via `xl_col_spec()` are left as-is.

Columns can be targeted by name or position. A plain data frame passed
alongside an `xl_sheet` behaves exactly as before:

```{r}
path <- write_xlsx(list(
  Styled = sheet,
  Raw    = data.frame(a = 1:3, b = letters[1:3])
), tempfile(fileext = ".xlsx"))
```

## Workbook defaults and metadata

`xl_workbook()` binds sheets to `xl_properties()`, which holds document
metadata plus the formatting defaults that would otherwise be built in. Because
those defaults are ordinary `xl_format` objects, you can change them — for
example, restyle the header row or set a workbook-wide default font.

```{r}
wb <- xl_workbook(
  list(Report = data.frame(quarter = c("Q1", "Q2"), sales = c(12000, 15000))),
  properties = xl_properties(
    title  = "Quarterly Report",
    author = "Finance Team",
    company = "Example Corp",
    custom = list(Project = "Alpha", Reviewed = TRUE),
    default_format = xl_font(name = "Calibri"),
    header_format  = xl_font(bold = TRUE, color = "white") +
                     xl_fill(background = "navy"),
    read_only = TRUE
  )
)
path <- write_xlsx(wb, tempfile(fileext = ".xlsx"))
```

Formats cascade from the workbook down: the effective format for a cell is
`default_format` merged with the column/row/header/hyperlink default, then with
the cell's own format. Identical resulting formats are written once, so
applying the same format to thousands of cells stays compact.

Note that a workbook-wide `default_format` is *emulated* (it is applied beneath
every cell) because the underlying library has no native "Normal style"
setter; it does not change Excel's built-in Normal style.

## Comments

Cell comments (notes) are attached through `xl_cell_general()`'s `comment`
argument. The simplest form is just text — a character vector, one comment per
cell (`NA` for none):

```{r}
df <- data.frame(item = c("Widget", "Gadget", "Gizmo"))
df$item <- xl_cell_general(
  value   = c("Widget", "Gadget", "Gizmo"),
  comment = c("Discontinued", NA, "Low stock")
)
path <- write_xlsx(df, tempfile(fileext = ".xlsx"))
```

`xl_comment()` adds options — an author, initial visibility, and box
size/position — and styling. Excel comment boxes support only a background
color and a font name/size/family, so styling is supplied by reusing the format
engine via `format`; any other format property triggers a warning.

```{r}
note <- xl_comment(
  "Please verify with Finance",
  author  = "Reviewer",
  visible = TRUE,
  format  = xl_font(name = "Arial", size = 10) + xl_fill(background = "lightyellow")
)
d <- data.frame(x = 1)
d$x <- xl_cell_general(value = 1234.5, comment = note)
path <- write_xlsx(d, tempfile(fileext = ".xlsx"))
```

A default author, and showing all of a sheet's comments at once, are set on
`xl_sheet()` (comment defaults are worksheet-scoped in Excel). A comment's own
`author` overrides the sheet default, and `visible` overrides `show_comments`:

```{r}
sheet <- xl_sheet(df, comment_author = "QA", show_comments = TRUE)
path <- write_xlsx(list(Data = sheet), tempfile(fileext = ".xlsx"))
```

## Page setup and printing

`xl_page_setup()` collects everything on Excel's "Page Layout" ribbon. Pass it
as `xl_sheet(page = )`. None of it touches the cell data --- it changes only
how the sheet prints and how it looks in page-break preview.

```{r}
sheet <- xl_sheet(df, page = xl_page_setup(
  orientation = "landscape",
  paper       = "A4",
  margins     = c(left = 0.5, right = 0.5),
  fit_to      = c(width = 1, height = 0),   # one page wide, any number tall
  header      = "&LQuarterly report&RPage &P of &N"
))
path <- write_xlsx(list(Data = sheet), tempfile(fileext = ".xlsx"))
```

`margins` takes one value for all four sides, four values in the order left,
right, top, bottom, or a named vector naming only the sides you care about.
`paper` takes a friendly name or an Excel paper-type integer for the envelope
and specialist sizes.

`fit_to` scales the printout to a number of pages, where `0` means "as many as
needed in that direction". `scale` sets a plain percentage instead; Excel
ignores it when `fit_to` is set.

### Headers and footers

Header and footer text uses Excel's own codes: `&L`, `&C` and `&R` start the
left, centre and right sections, `&P` is the page number, `&N` the page count,
`&D` the date, `&A` the sheet name, and `&&` a literal ampersand. The whole
string is limited to 255 characters.

```{r}
xl_page_setup(header = "&LDraft&CQ1&R&D", footer = "&CPage &P of &N")
```

Header and footer *images* are not supported yet, so an `&G` or `&[Picture]`
placeholder is rejected.

### Repeating headings and choosing what prints

```{r}
sheet <- xl_sheet(df, page = xl_page_setup(
  print_area  = "A1:D50",
  repeat_rows = 1,          # repeat the header row on every page
  repeat_cols = "A:A"       # and the first column
))
```

`print_area` takes any range writexl accepts, including a
`list(rows = , cols = )` spec. `repeat_rows` and `repeat_cols` take either a
count or a range string.

Note that these count **sheet** rows from 1, with the header row included ---
unlike `xl_row_spec()`, which indexes data rows. Repeating the header row is
the usual reason to use them, so it has to be nameable.

### Manual page breaks

`h_breaks` and `v_breaks` give 1-based positions at which a new page starts, so
`h_breaks = 21` breaks between rows 20 and 21.

```{r}
xl_page_setup(h_breaks = c(21, 41), v_breaks = "D")
```

Excel ignores manual breaks when `fit_to` is set, and writexl warns if you ask
for both.

## Sheet tabs and the opening view

`xl_sheet()` also controls the tab strip and what the user sees when the sheet
opens.

```{r}
wb <- list(
  Summary = xl_sheet(df, view = xl_sheet_view(active = TRUE, selection = "B2")),
  Detail  = xl_sheet(df, view = xl_sheet_view(hide_zero = TRUE)),
  Working = xl_sheet(df, view = xl_sheet_view(visible = FALSE))     # hidden tab
)
path <- write_xlsx(wb, tempfile(fileext = ".xlsx"))
```

`active` picks the tab Excel opens on, `selected` adds a tab to the selected
group, `visible = FALSE` hides one, and `first_tab` chooses the leftmost tab in
the strip.

These are the only worksheet settings whose rules span the whole workbook, and
writexl checks them before writing anything:

* a hidden sheet cannot be active or selected;
* at most one sheet may be active;
* the first sheet cannot be hidden unless another is made active, because Excel
  opens on the first sheet by default;
* at least one sheet must stay visible, or Excel refuses to open the file.

Each failure names the sheet at fault. libxlsxwriter enforces none of them, so
without these checks a bad combination would produce a workbook that simply
does not open.

### Where the sheet is scrolled to

`selection` sets the selected cell or range, and `top_left` the cell scrolled to
the top-left of the window:

```{r}
xl_sheet(df, view = xl_sheet_view(selection = "B2:D10", top_left = "A2"))
```

### Freezing and splitting

`freeze` locks rows and columns in place; `split` gives a movable divider with
independent scroll bars instead. They are mutually exclusive.

```{r}
xl_sheet(df, freeze = "A2")   # keep the header row visible
xl_sheet(df, view = xl_sheet_view(split  = "B3"))   # split above row 3 and left of column B
```

Note that libxlsxwriter positions a split by *distance*, in row-height and
column-width units, rather than by row and column — `15` means one default-height
row, `8.43` one default-width column. writexl converts the cell reference for
you using the sheet's actual row heights and column widths, so a split still
lands where you asked on a sheet whose rows or columns have been resized. Pass
`list(vertical = , horizontal = )` if you would rather give the units directly.

## Merged cells

`xl_merge()` merges a rectangle of cells into one, as Excel's "Merge and
Centre" does. Pass one or a list of them as `xl_sheet(merge = )`.

```{r}
sheet <- xl_sheet(df, merge = xl_merge(
  "A5:B5", "Total",
  format = xl_align(horizontal = "center") + xl_font(bold = TRUE)
))
path <- write_xlsx(list(Data = sheet), tempfile(fileext = ".xlsx"))
```

A merged range holds a single value, so `xl_merge()` carries its own `text`
rather than taking it from the data frame. Merges are applied after the sheet's
rows are written, so merging over cells the data frame filled keeps only the
merged text — exactly as merging in Excel discards everything but the top-left
value.

Excel has no single-cell merge, so a range covering one cell is an error.

Because a merge writes back over rows already emitted, any merge turns off the
memory-efficient row-streaming mode for the whole workbook. That is invisible
apart from a slightly different internal string storage, but it does mean the
whole sheet is held in memory while writing.
