---
title: "Formatting and workbook properties"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Formatting and workbook properties}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
library(writexl)
```

`writexl` writes plain data frames with sensible defaults, but you can also
control formatting at three levels that share one vocabulary:

* **cell** formatting, attached to columns with `xl_cell_general()`;
* **worksheet** layout, attached to a sheet with `xl_sheet()`;
* **workbook** defaults and metadata, attached with `xl_workbook()` /
  `xl_properties()`.

All three reuse the same [format engine](#the-format-engine), so once you know
how to build an `xl_format`, you can apply it anywhere.

## The format engine

A format is built from **group constructors**, each covering one family of
Excel properties. Every argument defaults to "unset", so you only specify what
you want to change.

```{r}
xl_font(bold = TRUE, color = "navy", size = 12)
xl_fill(background = "#FFF2CC")
xl_border(all = "thin", color = "gray")
xl_align(horizontal = "center", vertical = "top", wrap = TRUE)
xl_num_format("#,##0.00")
xl_protection(locked = FALSE)
```

Colors accept R color names (`"navy"`), hex strings (`"#FF0000"`), or integers,
via `xl_color()`.

Each group constructor already returns a complete `xl_format`, so a single
group can be used on its own. Combine groups with `+` (or `xl_format()`), which
merges **property by property** — later values win, but nothing you did not
touch is lost:

```{r}
fmt <- xl_font(bold = TRUE) + xl_fill(background = "yellow") + xl_num_format("0.0%")
fmt
```

## Cell formatting

Attach a format to a column of values with `xl_cell_general()`. Pass one format
for the whole column, or a list of formats for individual cells.

```{r}
df <- data.frame(item = c("Widget", "Gadget", "Gizmo"))
df$price <- xl_cell_general(
  value  = c(1234.5, 67.25, 890),
  format = xl_num_format("$#,##0.00") + xl_font(color = "darkgreen")
)
df$flag <- xl_cell_general(
  value  = c(10, -5, 3),
  format = list(
    xl_fill(background = "lightgreen"),
    xl_fill(background = "salmon"),
    xl_fill(background = "lightgreen")
  )
)
path <- write_xlsx(df, tempfile(fileext = ".xlsx"))
```

`xl_formula()` and `xl_hyperlink_cell()` also take a `format` argument. When a
cell holds a date or time and its format sets no number format, the default
date/time format is applied automatically.

## Worksheet layout

`xl_sheet()` wraps a data frame with column and row specifications and
sheet-level options. `xl_col_spec()` and `xl_row_spec()` are themselves
`xl_format` objects (so they combine with `+`) that additionally carry a target
and geometry.

```{r}
sheet <- xl_sheet(
  data.frame(
    date    = as.Date("2024-01-01") + 0:2,
    revenue = c(1000.5, 2000.25, 1500.75)
  ),
  cols = list(
    xl_col_spec("date", width = 12),
    xl_col_spec("revenue", width = 14,
                format = xl_num_format("#,##0.00") + xl_font(bold = TRUE))
  ),
  rows      = xl_row_spec(1, height = 20),
  freeze    = "A2",          # freeze the header row
  tab_color = "steelblue",
  zoom      = 120
)
path <- write_xlsx(list(Sales = sheet), tempfile(fileext = ".xlsx"))
```

`xl_sheet()` also adds **autofilters** and **worksheet protection**:

```{r}
locked <- xl_sheet(
  data.frame(id = 1:3, note = c("a", "b", "c")),
  autofilter = TRUE,                       # or a range like "A1:B4"
  protect    = list(password = "secret",   # protect, but still allow sorting
                    sort = TRUE)
)
path <- write_xlsx(list(Locked = locked), tempfile(fileext = ".xlsx"))
```

Cell locking set with `xl_protection()` only takes effect once the sheet is
protected.

Columns can be targeted by name or position. A plain data frame passed
alongside an `xl_sheet` behaves exactly as before:

```{r}
path <- write_xlsx(list(
  Styled = sheet,
  Raw    = data.frame(a = 1:3, b = letters[1:3])
), tempfile(fileext = ".xlsx"))
```

## Workbook defaults and metadata

`xl_workbook()` binds sheets to `xl_properties()`, which holds document
metadata plus the formatting defaults that would otherwise be built in. Because
those defaults are ordinary `xl_format` objects, you can change them — for
example, restyle the header row or set a workbook-wide default font.

```{r}
wb <- xl_workbook(
  list(Report = data.frame(quarter = c("Q1", "Q2"), sales = c(12000, 15000))),
  properties = xl_properties(
    title  = "Quarterly Report",
    author = "Finance Team",
    company = "Example Corp",
    custom = list(Project = "Alpha", Reviewed = TRUE),
    default_format = xl_font(name = "Calibri"),
    header_format  = xl_font(bold = TRUE, color = "white") +
                     xl_fill(background = "navy"),
    read_only = TRUE
  )
)
path <- write_xlsx(wb, tempfile(fileext = ".xlsx"))
```

Formats cascade from the workbook down: the effective format for a cell is
`default_format` merged with the column/row/header/hyperlink default, then with
the cell's own format. Identical resulting formats are written once, so
applying the same format to thousands of cells stays compact.

Note that a workbook-wide `default_format` is *emulated* (it is applied beneath
every cell) because the underlying library has no native "Normal style"
setter; it does not change Excel's built-in Normal style.
